---
title: Using Python sympy in a Python code chunk in RMarkdown
author: JD Long
date: '2020-07-27'
slug: using-python-sympy-in-a-python-code-chunk-in-rmarkdown
categories:
  - Python
tags:
  - python
  - rmarkdown
  - sympy
  - latex
---

I was recently asked about using `sympy` in Python to print $LaTeX$ formulas in RMarkdown notebooks. I misunderstood the question at first but after some thought I realized that I didn't know if it would work... but I should try it. 

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Yes, printing elements of a matrix. So it&#39;s due to the prog languages&#39; differences ( R vs Python), not rmarkdown vs Jupyter Notebook? <a href="https://t.co/Any4qIDHS7">pic.twitter.com/Any4qIDHS7</a></p>&mdash; Eyayaw T. Beze (@EyayawBeze) <a href="https://twitter.com/EyayawBeze/status/1287136564255105024?ref_src=twsrc%5Etfw">July 25, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

I knew that I could run Python chunks in an RMarkdown notebook. And I knew how to change the working conda environment so I could install `sympy`. But I wasn't so sure about getting the $LaTeX$ produced by `sympy` to print out properly. 

I tend to explicitly set my conda environment in my RMD file. Here I set it to `r-reticulate` which should be the default, but I do this so that it will remind me later what environment I'm using. Because when I come back to this in a few months I'll have forgotten. 

```{r}
library(reticulate)
use_condaenv(condaenv = "r-reticulate", required=TRUE)
```

to install conda packages from R or from your Rmd file you can do something like:

```{r eval=FALSE, include=TRUE}
reticulate::conda_install("r-reticulate","sympy")
```

Or you can use conda to activate the environment and install `sympy` with `conda install -c anaconda sympy`

The we can do magic like this:

Then we can run some Python chunks that depend on `sympy`
```{python, results='asis'}
from sympy import *
x = symbols('x')
a = Integral(cos(x)*exp(x), x)
output = Eq(a, a.doit()) 
print('$' + latex(output) + '$')
```

This works because of two pieces of code. One of which you can't see above. In the code chunk that runs above I set the language to python and told it to print out the results `asis`:

`{python, results='asis'}`

The `asis` flag tells `knitr` not to molest the resulting text string, just send it straight to the display. When I did that at first I ended up with this:

```{python, results='asis'}
from sympy import *
x = symbols('x')
a = Integral(cos(x)*exp(x), x)
output = Eq(a, a.doit()) 
print( latex(output)  )
```

Which baffled me because that's $LaTeX$ but it's not rendered... then (with some google help) I realized that in RMarkdown if we have stand alone $LaTeX$ we wrap them in dollar signs to tell `knitr` to render them... in this case using MathJax. So I just added `$` around the content like this:

```
print('$' + latex(output) + '$')
```

Of course if I were going to do this more than once I'd make a function to do it like this:

```{python}
from sympy import *
def latex_rmd_print(input_string):
   print('$' + latex(input_string) + '$')
```

Which then makes the earlier example a lot simpler... but don't forget the `results='asis'` in any code chunk that calls `latex_rmd_print`:

```{python, results='asis'}
x = symbols('x')
a = Integral( cos(x)*exp(x), x )
latex_rmd_print( Eq(a, a.doit()) )
```

